---
title: "East High Cycling Routes"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float:
      collapsed: false
      smooth_scroll: true
editor_options:
  chunk_output_type: console
---

# Input Data & Configuration

## Libraries

```{r libs, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = TRUE, message = FALSE}
date()
rm(list=ls())
library(tidyverse)
library(ggmap)
library(sf)
library(osrm)
library(smoothr)
library(magick)
library(ggnewscale)
library(rsvg)
library(httr)
library(jsonlite)
library(parallel)
fig.height <- 6
set.seed(1)
runLTS <- FALSE
logo <- image_read(path = "other/BFW_Logo_180_x_200_transparent_background.png")
school_symbol <- image_read_svg(path = "other/school_FILL0_wght400_GRAD0_opsz24.svg")
```

## School Location Data

```{r gpkg, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = TRUE, message = FALSE}
WI_schools <- st_transform(st_read(dsn = "data/Schools/Wisconsin_Public_Schools_-5986231931870160084.gpkg"), crs = 4326)
WI_schools <- WI_schools %>% mutate(geom = SHAPE)
```

## Addresses Data

```{r addresses, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = TRUE, message = FALSE}
addresses <- read_csv(file="data/addresses/Addresses_Students_EastHS_2024_GeocodeResults.csv") %>%
  filter(lat > 0) %>%
  st_as_sf(coords=c("lon","lat"), crs=4326)
```
(Remember that x = lon and y = lat.)

## Bike Level of Traffic Stress (LTS)

```{r bikelts, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = TRUE, message = FALSE}
bike_lts <- st_transform(st_read("data/bike_lts/bike_lts_DANE.geojson"), crs = 4326)
# make lts attribute a factor
bike_lts[["lts"]] <- as.factor(bike_lts$LTS_F)
# remove segments with an LTS value of 9
bike_lts <- bike_lts %>% filter(lts != 9)

# set color scale
bike_lts_scale <- data.frame(code = c(1, 2, 3, 4, 9),
                             color = c("#1a9641",
                                       "#a6d96a",
                                       "#fdae61",
                                       "#d7191c",
                                       "#d7191c"))
```

# External sources configurations

## Open Source Routing Machine (OSRM)

```{r osrm, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = TRUE, message = FALSE}
# Set url and profile of OSRM server
options(osrm.server = "http://127.0.0.1:5001/")
options(osrm.profile = "bike")
```

## Brouter options
```{r brouter, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = TRUE, message = FALSE}
# Set url and profile of brouter server
brouter_url <- "http://127.0.0.1:17777/brouter"
brouter_profile <- "safety"
```

## Stadia Maps API Key

```{r stadiamaps, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = TRUE, message = FALSE}
register_stadiamaps(key = substr(read_file(file = "api_keys/stadia_api_key"), 1, 36))
```
# Analysis

```{r analysisPreamble, eval = TRUE, echo = FALSE, results = "show", warning = FALSE, error = TRUE, message = FALSE}
radius <- 3 # miles
levels <- c(1)
res <- 100
threshold <- 1
```

## Subset Addresses Within `r radius` Miles

```{r cycleBoundary, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = TRUE, message = FALSE}
cycle_boundary_m <- radius*1609
school_focus <- data.frame(name = c("East High School"), NCES_CODE = c("550852000925"))
#school_focus <- data.frame(name = c("IMAP"), NCES_CODE = c("550008203085"))

cycle_boundary_poly <- st_transform(fill_holes(st_make_valid(osrmIsodistance(
  loc = WI_schools %>% filter(NCES_CODE %in% school_focus$NCES_CODE),
#  breaks = c(cycle_boundary_m),
  breaks = cycle_boundary_m*levels,
  res = res)
), units::set_units(threshold, km^2)), crs = 4326)

addresses_near <- st_intersection(addresses, cycle_boundary_poly)
```
Notes:

- _osrmIsoDistance_ is the primary function in the above chunk.
- This function computes areas that are reachable within a given road
distance from a point and returns the reachable regions as
polygons. These areas of equal travel distance are called isodistances.
- Input is a point represented as an sf object (extended
data.frame-like objects with a simple feature list column) could be
other classes, e.g., vector of coods, data.frame of lat tand
long. etc.
- Arguments to osrmIsodistances used here are breaks and res
 - breaks: a numeric vector of break values to define isodistance areas, in meters.
 - res: number of points used to compute isodistances, one side of the
square grid, the total number of points will be res*res. Increase res to obtain more detailed isodistances.
- _fill\_holes_ is also used with a threshold of `r threshold` km^2.
- _st\_intersection_ is also used on sf objects (simple features?)

## Calculate Routes

```{r routes, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = TRUE, message = FALSE}
routes <- list(NULL)
school_focus_location <- WI_schools %>% filter(NCES_CODE %in% school_focus$NCES_CODE) %>% select(LAT, LON)
for(i in addresses_near %>% arrange(number) %>% pull(number)) {
  query <- paste0(
    brouter_url,
    "?lonlats=",
    (addresses_near %>% filter(number == i) %>% pull(point) %>% str_split(., ","))[[1]][1], ",",
    (addresses_near %>% filter(number == i) %>% pull(point) %>% str_split(., ","))[[1]][2], "|",
    school_focus_location$LON, ",", school_focus_location$LAT,
    "&profile=", brouter_profile,
    "&alternativeidx=0&format=geojson"
  )
  response <- GET(query)
  route_run <- st_read(content <- content(response, as = "text"), quiet = TRUE)
  route_run[["student_number"]] <- i
  routes[[i]] <- route_run


  message(paste0("done - ", i, " of ", max(addresses_near$number)))
}

routes <- st_transform(bind_rows(routes), crs = 4326)
```

Notes:
- this queries the brouter server to get routes

## Set boundaries and get basemap
```{r basemap, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = TRUE, message = FALSE}

bbox <- st_bbox(st_buffer(cycle_boundary_poly, dist = 500))
bbox <- c(left = as.double(bbox[1]),
          bottom = as.double(bbox[2]),
          right = as.double(bbox[3]),
          top = as.double(bbox[4]))

#get basemap
basemap <- get_stadiamap(bbox = bbox, zoom = 15, maptype = "stamen_toner_lite")
```
Notes:
- This chunk retrieves the base map from Stadia Maps (API key required)


## Combine routes with Bike LTS
```{r ltscount, eval = runLTS, echo = TRUE, results = "show", warning = FALSE, error = TRUE, message = FALSE}

# Count the routes that intersect or overlap with each segment of the bike_tls network.
# The intersections have a buffer of 10m
bike_lts_buffer <- st_buffer(st_intersection(bike_lts, cycle_boundary_poly), 10)

bike_lts_buffer["student_use"] <- unlist(lapply(st_intersects(bike_lts_buffer, routes), length))

bike_lts <- left_join(bike_lts, as.data.frame(bike_lts_buffer %>% select(OBJECTID, student_use)), by = "OBJECTID")
```

Notes:
- for each segment in bike_lts, this counts how many student&rsquo;s
  calculated routes intersect with it (within a 10 m buffer)

```{r functions, eval = runLTS, echo = FALSE, results = "show", warning = FALSE, error = TRUE, message = FALSE}
source("./R/functions.R")
```

```{r routeslts, eval = runLTS, echo = TRUE, results = "show", warning = FALSE, error = TRUE, message = FALSE}
# Start with routes_lts as a NULL list
routes_lts <- list(NULL)

# Pre-filter the bike_lts_buffer for relevant student use
relevant_buffer <- bike_lts_buffer %>% filter(student_use > 0)

routes_lts <- mclapply(addresses_near %>% arrange(number) %>% pull(number),
                       getLTSForRoute,
                       mc.cores = detectCores() / 2,
                       mc.cleanup = TRUE,
                       mc.preschedule = TRUE,
                       mc.silent = FALSE)

routes_lts <- bind_rows(routes_lts)
```
Notes:
- for each student's route, this finds which bike_lts segment it
  intersects with and calculates a max and an average level of traffic
  stress (LTS). This takes a while, so a parallelized it. There's
  probably a more efficient way to do this calculation.
- see ./R/functions.R for defintion of getLTSForRoute()


# Make Maps

## Generate map with LTS data

```{r maplts, eval = runLTS, echo = FALSE, results = "show", warning = FALSE, error = TRUE, message = FALSE}
ggmap(basemap) +
  geom_sf(data = routes_lts %>% filter(student_number == 6), inherit.aes = FALSE,
          aes(color = route$lts,
              geometry = route$geometry),
          linewidth = 2)  +
  scale_color_manual(values = bike_lts_scale$color, name = "Bike Level of Traffic Stress")

# Join the data with the addresses data
addresses_near <- left_join(addresses_near,
                            routes_lts %>%
                              select(c("student_number", "lts_max", "lts_average", "lts_1_dist", "lts_2_dist", "lts_3_dist", "lts_4_dist")),
                            join_by("number"=="student_number"),
                            multiple = "any")

# add supplemental analysis
addresses_near <- addresses_near %>% mutate(lts_34_dist = lts_3_dist + lts_4_dist)
```

## Generate map of addresses
```{r mapaddresses, eval = TRUE, echo = FALSE, results = "show", warning = FALSE, error = TRUE, message = FALSE}

ggmap(basemap) +
  labs(title = paste0("Student homes at ",
                      school_focus %>% pull(name)),
       x = NULL,
       y = NULL,
       color = NULL,
       fill = "How many students live there") +
  theme(axis.text=element_blank(),
        axis.ticks=element_blank(),
        plot.caption = element_text(color = "grey")) +
  geom_hex(data = addresses %>% extract(geometry, into = c('Lat', 'Lon'), '\\((.*),(.*)\\)', conv = T),
           aes(x = Lat,
               y = Lon),
           alpha = 0.7) +
  scale_fill_distiller(palette = "YlOrRd", direction = "reverse") +
  geom_sf(data = cycle_boundary_poly,
          inherit.aes = FALSE,
          aes(color = paste0(radius, " mile cycling boundary")),
          fill = NA,
          linewidth = 1) +
  scale_color_manual(values = "blue", name = NULL) +
  new_scale_color() +
  annotation_raster(school_symbol,
                    # Position adjustments here using plot_box$max/min/range
                    ymin = as.double((WI_schools %>% filter(NCES_CODE %in% school_focus$NCES_CODE) %>% pull(geom))[[1]])[2] - 0.001,
                    ymax = as.double((WI_schools %>% filter(NCES_CODE %in% school_focus$NCES_CODE) %>% pull(geom))[[1]])[2] + 0.001,
                    xmin = as.double((WI_schools %>% filter(NCES_CODE %in% school_focus$NCES_CODE) %>% pull(geom))[[1]])[1] - 0.0015,
                    xmax = as.double((WI_schools %>% filter(NCES_CODE %in% school_focus$NCES_CODE) %>% pull(geom))[[1]])[1] + 0.0015) +
  geom_sf_label(data = WI_schools %>% filter(NCES_CODE %in% school_focus$NCES_CODE),
                inherit.aes = FALSE,
                mapping = aes(label = school_focus %>% pull(name)),
                nudge_y = 0.0015,
                label.size = 0.04,
                size = 2)
ggsave(file = paste0("figures/",
                     school_focus %>% pull(name),
                     " Addresses_cycling.pdf"),
       title = paste0(school_focus %>% pull(name), " Addresses"),
       device = pdf,
       height = 8.5,
       width = 11,
       units = "in",
       create.dir = TRUE)
```

## Generate map of routes
```{r maproutes, eval = TRUE, echo = FALSE, results = "show", warning = FALSE, error = TRUE, message = FALSE}
# generate map
ggmap(basemap) +
  labs(title = paste0("Cycling routes for students at ",
    school_focus %>% pull(name)),
    subtitle = paste0("only showing routes within the ", radius, " mile cycling boundary"),
    x = NULL,
    y = NULL,
    color = NULL,
    linewidth = "Potential student cyclists") +
  theme(axis.text=element_blank(),
        axis.ticks=element_blank(),
        plot.caption = element_text(color = "grey")) +
  geom_sf(data = cycle_boundary_poly,
          inherit.aes = FALSE,
          aes(color = paste0(radius, " mile cycling boundary")),
          fill = NA,
          linewidth = 1) +
  scale_color_manual(values = "blue", name = NULL) +
  new_scale_color() +
  geom_sf(data = bike_lts %>% filter(!is.na(student_use), student_use > 3),
          inherit.aes = FALSE,
          aes(linewidth = student_use),
          color = "mediumvioletred",
          fill = NA) +
  scale_linewidth_continuous(range = c(0, 3)) +
  annotation_raster(school_symbol,
                    # Position adjustments here using plot_box$max/min/range
                    ymin = as.double((WI_schools %>% filter(NCES_CODE %in% school_focus$NCES_CODE) %>% pull(geom))[[1]])[2] - 0.001,
                    ymax = as.double((WI_schools %>% filter(NCES_CODE %in% school_focus$NCES_CODE) %>% pull(geom))[[1]])[2] + 0.001,
                    xmin = as.double((WI_schools %>% filter(NCES_CODE %in% school_focus$NCES_CODE) %>% pull(geom))[[1]])[1] - 0.0015,
                    xmax = as.double((WI_schools %>% filter(NCES_CODE %in% school_focus$NCES_CODE) %>% pull(geom))[[1]])[1] + 0.0015) +
  geom_sf_label(data = WI_schools %>% filter(NCES_CODE %in% school_focus$NCES_CODE),
                inherit.aes = FALSE,
                mapping = aes(label = school_focus %>% pull(name)),
                nudge_y = 0.0015,
                label.size = 0.04,
                size = 2)

ggsave(file = paste0("figures/",
                     school_focus %>% pull(name),
                     " Routes_cycling.pdf"),
       title = paste0(school_focus %>% pull(name), " Cycling Routes"),
       device = pdf,
       height = 8.5,
       width = 11,
       units = "in",
       create.dir = TRUE)
```

## Generate map of routes with LTS (1)
```{r maprouteslts, eval = runLTS, echo = FALSE, results = "show", warning = FALSE, error = TRUE, message = FALSE}
# generate map
ggmap(basemap) +
  labs(title = paste0("Cycling routes for students at ",
                      school_focus %>% pull(name)),
       subtitle = "only showing routes within the cycling boundary",
       x = NULL,
       y = NULL,
       color = NULL,
       linewidth = "Potential student cyclists") +
  theme(axis.text=element_blank(),
        axis.ticks=element_blank(),
        plot.caption = element_text(color = "grey")) +
  geom_sf(data = cycle_boundary_poly,
          inherit.aes = FALSE,
          aes(color = paste0(radius, " mile cycling boundary")),
          fill = NA,
          linewidth = 1) +
  scale_color_manual(values = "blue", name = NULL) +
  new_scale_color() +
  geom_sf(data = bike_lts %>% filter(!is.na(student_use), student_use > 0),
         inherit.aes = FALSE,
         aes(color = lts,
             linewidth = student_use)) +
  scale_color_manual(values = bike_lts_scale$color, name = "Bike Level of Traffic Stress") +
  scale_linewidth_continuous(range = c(0, 3)) +
  annotation_raster(school_symbol,
                    # Position adjustments here using plot_box$max/min/range
                    ymin = as.double((WI_schools %>% filter(NCES_CODE %in% school_focus$NCES_CODE) %>% pull(geom))[[1]])[2] - 0.001,
                    ymax = as.double((WI_schools %>% filter(NCES_CODE %in% school_focus$NCES_CODE) %>% pull(geom))[[1]])[2] + 0.001,
                    xmin = as.double((WI_schools %>% filter(NCES_CODE %in% school_focus$NCES_CODE) %>% pull(geom))[[1]])[1] - 0.0015,
                    xmax = as.double((WI_schools %>% filter(NCES_CODE %in% school_focus$NCES_CODE) %>% pull(geom))[[1]])[1] + 0.0015) +
  geom_sf_label(data = WI_schools %>% filter(NCES_CODE %in% school_focus$NCES_CODE),
                inherit.aes = FALSE,
                mapping = aes(label = school_focus %>% pull(name)),
                nudge_y = 0.0015,
                label.size = 0.04,
                size = 2)

ggsave(file = paste0("figures/",
                     school_focus %>% pull(name),
                     " Routes - Traffic Stress_cycling.pdf"),
       title = paste0(school_focus %>% pull(name), " Cycling Routes - Traffic Stress"),
       device = pdf,
       height = 8.5,
       width = 11,
       units = "in",
       create.dir = TRUE)

```

## Generate map of routes with LTS (2)

```{r mapaddresseslts, eval = runLTS, echo = FALSE, results = "show", warning = FALSE, error = TRUE, message = FALSE}
# generate map
ggmap(basemap) +
  labs(title = paste0("Level of Traffic stress for biking for students at ",
                      school_focus %>% pull(name)),
       subtitle = "only showing routes within the cycling boundary",
       x = NULL,
       y = NULL) +
  theme(axis.text=element_blank(),
        axis.ticks=element_blank(),
        plot.caption = element_text(color = "grey")) +
  geom_sf(data = cycle_boundary_poly,
          inherit.aes = FALSE,
          aes(color = paste0(radius, " mile cycling boundary")),
          fill = NA,
          linewidth = 1) +
  scale_color_manual(values = "blue", name = NULL) +
  new_scale_color() +
  geom_sf(data = routes_lts %>% filter(route$student_use >= 4),
         inherit.aes = FALSE,
         aes(geometry = route$geometry,
             color = route$lts,
             linewidth = route$student_use)) +
  #scale_color_gradientn(colors = bike_lts_scale$color, name = "Length of high stress travel on route from that address", limits = c(1,4)) +
  scale_color_manual(values = bike_lts_scale$color, name = "Bike Level of Traffic Stress") +
  #scale_color_distiller(palette = "YlOrRd", direction = "reverse") +
  scale_linewidth_continuous(range = c(0, 3)) +
  annotation_raster(school_symbol,
                    # Position adjustments here using plot_box$max/min/range
                    ymin = as.double((WI_schools %>% filter(NCES_CODE %in% school_focus$NCES_CODE) %>% pull(geom))[[1]])[2] - 0.001,
                    ymax = as.double((WI_schools %>% filter(NCES_CODE %in% school_focus$NCES_CODE) %>% pull(geom))[[1]])[2] + 0.001,
                    xmin = as.double((WI_schools %>% filter(NCES_CODE %in% school_focus$NCES_CODE) %>% pull(geom))[[1]])[1] - 0.0015,
                    xmax = as.double((WI_schools %>% filter(NCES_CODE %in% school_focus$NCES_CODE) %>% pull(geom))[[1]])[1] + 0.0015) +
  geom_sf_label(data = WI_schools %>% filter(NCES_CODE %in% school_focus$NCES_CODE),
                inherit.aes = FALSE,
                mapping = aes(label = school_focus %>% pull(name)),
                nudge_y = 0.0015,
                label.size = 0.04,
                size = 2)

ggsave(file = paste0("figures/",
                     school_focus %>% pull(name),
                     " Routes - Traffic Stress_cycling_new.pdf"),
       title = paste0(school_focus %>% pull(name), " Student Addresses - Cycling Traffic Stress"),
       device = pdf,
       height = 8.5,
       width = 11,
       units = "in",
       create.dir = TRUE)

```

# Appendix

## Notes

### R Package sf

- Simple features or simple feature access refers to a formal standard
  (ISO 19125-1:2004) that describes how objects in the real world can
  be represented in computers, with emphasis on the spatial geometry
  of these objects. It also describes how such objects can be stored
  in and retrieved from databases, and which geometrical operations
  should be defined for them.
- The standard is widely implemented in spatial databases (such as
  PostGIS), commercial GIS (e.g., ESRI ArcGIS) and forms the vector
  data basis for libraries such as GDAL. A subset of simple features
  forms the GeoJSON standard.
- R has well-supported classes for storing spatial data (sp) and
  interfacing to the above mentioned environments (rgdal, rgeos), but
  has so far lacked a complete implementation of simple features,
  making conversions at times convoluted, inefficient or
  incomplete. The package sf tries to fill this gap, and aims at
  succeeding sp in the long term.
- See source [here](https://r-spatial.github.io/sf/articles/sf1.html)
- all functions and methods in sf that operate on spatial data are
  prefixed by st_, which refers to spatial type; this makes them
  easily findable by command-line completion.
- Tessellation st_make_grid()

### Sandbox

#### Create a Grid Over Bikeable Area

```{r sandbox1, eval = TRUE, echo = TRUE, results = "show", warning = TRUE, error = TRUE, message = TRUE}
cellsize <- 5e-3

grid <- st_intersection(cycle_boundary_poly, st_make_grid(cycle_boundary_poly, cellsize = cellsize, what = "polygons", square = FALSE))

ggmap(basemap) +
  geom_sf(data = st_geometry(grid),
          inherit.aes = FALSE,
          fill = NA,
          linewidth = 0.5)
```

#### Compute Routes from Centroid to School

```{r sandbox2, eval = TRUE, echo = TRUE, results = "show", warning = TRUE, error = TRUE, message = TRUE}
grid_pts <- st_centroid(grid)
grid_coods <- st_coordinates(grid_pts)
routes <- list(NULL)

for(i in 1:nrow(grid_coods) ) {
  query <- paste0(
    brouter_url,
    "?lonlats=", grid_coods[i,1], ",",grid_coods[i,2], "|",
    school_focus_location$LON, ",", school_focus_location$LAT,
    "&profile=", brouter_profile,
    "&alternativeidx=0&format=geojson"
  )
  response <- GET(query)
  route_run <- st_read(content <- content(response, as = "text"), quiet = TRUE)
  route_run[["student_number"]] <- i
  routes[[i]] <- route_run
}

routes <- st_transform(bind_rows(routes), crs = 4326)

ggmap(basemap) +
  geom_sf(data = st_geometry(routes),
          inherit.aes = FALSE,
          fill = NA,
          linewidth = 0.5,
          color = "red"
          ) +
      geom_sf(data = st_geometry(grid),
          inherit.aes = FALSE,
          fill = NA,
          linewidth = 0.5)
```

#### Compute Route-Level Measures

How do we retireve features of the route such as percentage on
cycleway? My next step is to look over getLTSForRoute() for tips.

```{r chunklast, eval = TRUE, echo = TRUE, results = "show", warning = TRUE, error = TRUE, message = TRUE}
date()
sessionInfo()
```

# Archive

```{r archive1, eval = FALSE, echo = TRUE, results = "show", warning = TRUE, error = TRUE, message = TRUE}
# for(i in addresses_near %>% arrange(number) %>% pull(number)) {
#   lts_segments <- bike_lts_buffer$OBJECTID[st_intersects(bike_lts_buffer, routes %>% filter(student_number == i), sparse = FALSE)]
#   lts_max <- max(bike_lts_buffer %>% filter(OBJECTID %in% lts_segments) %>% pull(LTS_F), na.rm = TRUE)
#   lts_average <- mean(bike_lts_buffer %>% filter(OBJECTID %in% lts_segments) %>% pull(LTS_F), na.rm = TRUE)
#   routes_lts[[i]] <- data.frame("student_number" = c(i), "lts_max" = c(lts_max), "lts_average" = c(lts_average))
#   message(paste0("done - ", i, " of ", max(addresses_near$number)))
# }

# routes_lts <- lapply(head(addresses_near %>% arrange(number) %>% pull(number)),
#      getLTSForRoute)

# system.time(routes_lts <- lapply(head(addresses_near %>% arrange(number) %>% pull(number)),
#       getLTSForRoute))

```
